---
title: 'Reuniting Accessibility Measures with Spatial Interaction Principles'
author: 'Anastasia Soukhov'
date: 'July 2025'
university: 'McMaster University'
faculty: 'Faculty of Science'
department: 'School of Earth, Environment and Society'
subject: 'Geography'
supervisor: 'Antonio Páez'
bdegree: 'B.Eng.' # Bachellor's Degree
mdegree: 'M.A.Sc.'   # Master's Degree (if any)
# This is the degree you are aiming for with this thesis
degree: 'Doctor of Philosophy'
knit: bookdown::render_book
site: bookdown::bookdown_site
output:
  macdown::thesis_pdf: default 
  keep_tex: true
lay_abstract: |
  One of the primary goals of transportation systems is to connect people with opportunities--such as jobs, essential services and community. However, traditional transportation planning practice has largely emphasized mobility, focusing on metrics like distance traveled or travel time, rather than accessibility (the number of opportunities people can reach). Although researchers have long advocated for a shift from mobility-based approaches to accessibility-focused ones, this transition has yet to be fully realized due to several persistent challenges. A key barrier is methodological: the absence of standardized units for measuring accessibility. This thesis seeks to address this barrier by examining how accessibility methods relate to conventional mobility-based techniques, and by exploring how constraints--adapted from spatial interaction models--can be used to return meaningful units to accessibility values and in accessibility analysis. This is done by reviewing the related spatial interaction modeling and accessibility literature, outlining a family of 'constrained' accessibility measures, and detailing an empirical example of accessible parkland area and population in the City of Toronto.
  
abstract: |
  Transportation systems play a fundamental role in facilitating access between people and opportunities across place. Place-based accessibility can be defined as the potential of opportunities for spatial interaction. For decades, transportation planning research and practice have relied on mobility-based metrics, such as kilometers traveled, emissions released, or vehicle counts--measures that reflect realised movement. In contrast, accessibility focuses on potential movement, such as the number of opportunities that can be reached within given constraints. Although there has been growing interest in shifting from mobility-based methods to access-focused planning approaches, a few barriers remain. One significant issue is methodological--namely, the lack of clarity in how accessibility values are interpreted. There are several approaches to match values with meaning, but this dissertation proposes a preceding step: clarifying the units of accessibility.

  In response, this dissertation is organized into six chapters. First, a review of how accessibility literature largely diverged from the spatial interaction literature is detailed. This lays the foundation for how the addition of proportionality constants that return the units to the measure and balance them to reflect known constraints in the system may be useful. Second, using a synthetic example, the total constrained accessibility measure is introduced. This measure is formally proportional to unconstrained accessibility (i.e., Hansen-type accessibility measure commonly used in literature) but with results consistently expressed in units of accessible opportunities or units of accessible population. Next, using the same synthetic example, the single constrained accessibility measure is introduced. This measure considers competition for opportunities and could be understood as related to the popular measures of competitive accessibility. Multimodal extensions as well as the expression of these measure in terms of ‘market potential’--the accessible population at a zone--is also introduced. Third, motivation, data and methods of an empirical example of the parkland area and population in the City of Toronto are detailed. All members of the family of accessibility measures introduced in this work are compared and contrasted in the subsequent chapters using this empirical example. Fourth, a unimodal walking-only case considering a binary impedance function and all population is presented, where the accessible parkland area and accessible population in Toronto is quantified. Differences in the unconstrained (i.e., conventional method) and constrained measures are detailed. Fifth, a multimodal case of accessible parkland area and accessible population is detailed, considering walking, cycling, transit, and car modes. Similarly, differences in the unconstrained and constrained measures are detailed. Sixth, this final chapter summarizes key points across all chapters, provides a thorough discussion on how constrained accessibility can be used to communicate for the purpose of equitable transportation policy, and details future lines of potential investigation.
  
# Note that a tab is needed on the line after the `|`. Subsequent paragraphs need to be indented too.
acknowledgements: |
  Oh the places this dissertation has taken me.
  
  First and foremost, I want to express gratitude to my PhD supervisor Antonio Páez, for your unwavering and steady guidance throughout this journey. From co-working coding and writing sessions and weekly check-ins to countless opportunities to present my work in regional and international venues: the freedom and support you've given me to develop ideas, my academic voice and establish my own network over the past four years have made the completition of this dissertation possible. I also want to sincerly thank Moataz Mohamed, my MASc supervisor, who first folded me into to the academic world. Without you, these ~180 pages would not be written. I want to also thank Christopher D. Higgins, who worked alongside Antonio, Moataz and myself on one of the earliest projects of my PhD journey--work that ultimately laid the foundation for this dissertation.

  I also want to express gratitude for the Mobilizing Justice Partnership, led by Steven Farber and supported academically in its early years by Ignacio Tiznado-Aitken and Matthew Palm--thank you for the guidance and support in framing my academic understanding of transportation equity and justice. I'm also especially grateful for the Partnerships'connection with the University of Toronto and the opportunity to work with others with similar synergies; João Pedro Figueira Amorim Parga, Robert Nutifafa Arku, Madeleine Bonsma-Fisher. Also, a major thank you to the Mobilizing Justice Partnership’s network of transportation practitioners committed to advancing equity. Through this network, I had the opportunity to engage with the City of Toronto’s work in developing the Transportation Equity Policy Framework, led by Lorina Hoxha. As a member of the technical advisory committee, I witnessed firsthand the dedication of public servants pushing the boundaries of defining equity analysis in practice. I was grateful to contribute feedback on methods in various forums, faciliate workshops and participate in some lively discussions with fellow accessibility fanatics--Michael Hain, Bryce Sherman, and Herman Hui. These conversations inspired me to apply the methods introduced in this dissertation to a practical example they were exploring. I extend my sincere gratitude to all of them.

  I would also like to extend my gratitude to those at McMaster University. I am particularly grateful to Léa Ravensbergen, whose mentorship opened the door to my first (and subsequent) co-supervisions of passionate undergraduate students at McMaster University: Nicholas Mooney, Isla Leadbetter, and Angel Emmanuel Hernandez Cruz. I also want to thank the growing community of transportation reseachers--the Friends of Transportation Fest--Darren Scott and his students Yehui Yin, Haoran Xu, and Sadia Chowdhury, along with my academic siblings Bruno Dos Santos and Amanda Rorat and visiting 'siblings' Javier Tarriño-Ortiz and Carlos Marigil Alba, along with the comraderie of Léa students: Alexandra Sbrocchi, Dastar Bhoday, Domenica De Iuliis, and Moataz's students particuarly: Sadia Tasnim, Ali Shehabeldeen and Ahmed Foda. Though I completed the PhD program and this dissertation while living in Toronto, you all made visiting campus worthwhile--rewarding, enriching, and fun. Your perspectives sharpened my thinking and helped me see familiar problems in new ways.

  To my international friends and collaborators, a big thank you! I’m especially grateful to Rafael H. M. Pereira was looped into my work during his stay as an visiting professor at UofT in Fall 2024 / Spring 2025. I'm also grateful to everyone who made my two spring research stays at TRANSyT at the Universidad Politécnica de Madrid in 2023 and 2024 productive, tremendously fun, and ultimately unforgettable. Thank you to Julio A. Soria-Lara, Alberto Rojas-Rivero, Manuel Pérez, Raúl F. Elizondo-Candanedo, Miguel L. Navarro-Ligero, Amor Ariza-Álvarez, and all the other lab mates at TRANSyT for your support, collaboration, and camaraderie. And to the friends, colleagues, and co-conspirators I've had the good fortunte of connecting with, gaining inspiration from, and recieving feedback from at international conferences (i.e., NARSC, AAG, TRB, WCTRS, and NECTAR): thank you. To the co-organizers of the Transportation Justice group at AAG and the members of the NECTAR Board--thank you for creating spaces that foster both a passion for advancing transportation research and a sense of academic community. These international connections made the pursuit of a PhD all the more worthwhile.

  Finally, to my partner in life--and now husband--Aidan: thank you for your patience, love, and unwavering belief in me and in humanity’s pursuit of knowledge for a better world. From encouraging me to take the plunge into the PhD program, to supporting my obsessive late-night working sessions before (often self-imposed) deadlines, to planning our wedding while I was furiously writing away on this dissertation, to standing by me through weeks and months apart during conferences and research stays--your support has been constant. Your belief in me, and in the value of continuing this academic journey, means more than I can ever fully express.

  This work stands on the shoulders of many: many more than I can name possibly list. But to all who came before me, thank you for charting the path I was lucky enough to find myself on and follow. Thank you for making it possible for me to inch just a bit further.

declarations: |
 I hereby declare that I am the primary author of this dissertation. This is a true copy of the dissertation, including any required final revisions.
 
 I authorize McMaster University to lend this disertations to other institutions or individuals for the purpose of scholarly research as well as reproduce this dissertation, in total or in part for the purpose of scholarly research.
 
 I understand that my MRP may be made electronically available to the public
  
quotation: 'There is no such thing as a new idea. It is impossible. We simply take a lot of old ideas and put them into a sort of mental kaleidoscope... they make new and curious combinations...' 

# Author of the quotation
qauthor: 'Mark Twain'
# Specify the location where the bibliography file is found below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include = FALSE}
# This chunk ensures that the macdown package is
# installed and loaded. This macdown package includes
# the template files for the thesis.
if(!require(remotes))
  install.packages("remotes", repos = "http://cran.rstudio.com")
if(!require(macdown))
  remotes::install_github("paezha/macdown")
library(macdown)
```

<!--
The acknowledgements, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
-->

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

## This dissertation's evolution

This dissertation has taken shape non-chronologically. In the spring of 2021, still a largely work-from-home period for many due to COVID-19 Pandemic public health measures, I began a reading course with Antonio Páez as a Masters of Applied Science student under the supervision of Moataz Mohammed. At that point, I was nearing the end of the Master's program, on the way to completing a MASc. thesis aimed at providing guidance on right-sizing passenger transportation sustainability policies based on vehicle technologies well-to-wheel life cycle operating conditions e.g., [@soukhovOccupancyGHGEmissions2022; @soukhovElectricMobilityEmission2022].

Coming from civil and transportation engineering, I had to take on a new research approach to engage with the project introduced to me by Antonio. Antonio, who had been developing a deep appreciation for open and reproducible science and the R universe, combined with a passion for use of data for advocacy, had worked with students before me to gather information on public schools in Hamilton. A public school in his neighbourhood, along with several others across the city, was closing, and he was keen to explore the impacts these closures would have. It was within this context that I approached this reading course.

I found myself intuitively working through the available data, building upon my beginner R skills, estimating additional data to supplement the analysis, and crafting research questions that felt sufficiently satisfying; questions that blended policy impacts, spatial analysis, and active transportation implications of a change in school-seats. One of these implications would be the change in the number of school-seats available to students, especially those who were within walking distance of a school that was closed. Place-based accessibility, or the potential to spatially interact with school-seats, was the indicator that could quantify this intuition. Accessibility, conventionally defined as the sum of opportunities $O_j$ weighted by the travel impedance function $f(c_{ij})$ of some travel cost $c_{ij}$ for a set of zones of origin $i$ $i...I$ and destination $j$ $j...J$ in Equation \ref{eq:access-unconstrained}, takes the following general form:

```{=tex}
\begin{equation}
\label{eq:access-unconstrained}
S_i = \sum_i O_j f(c_{ij}) 
\end{equation} 
```

For this reading course project, I wanted as precise a solution as possible, with clear interpretability and applicable at the highest spatial resolution. For instance: how had the number of potentially reachable school-seats change for households? Reflecting on this thought process, I was placing very high expectations on a place-based accessibility metric--an expectation shaped by my engineering background which played a role in driving my focus on precise and measurable units. For this project, I helped gather the number of school-seats in Hamilton for each study year, the number of students, and I knew what I wanted the measure to reflect my intuition: that the school-seat accessibility for each household should decrease within proximity to schools. However, conventional accessibility methods produced results that were not in 'the number of school-seats', thereby being more difficult to intuitively interpret. Moreover, competitive accessibility measures, such as those often applied using Floating Catchment Areas (FCA), had their own inflationary issues [i.e., discussed in @paezDemandLevelService2019]. 

So, a new method was needed to satisfy the quest set up for myself in this reading course. What we (myself, and ultimately co-authors Chris Higgins, Moataz Mohamed and Antonio Paez) ultimately came up with was an intuitivively formulated method using proportional allocation factors. These factors allocate opportunities (school-seats) proportionally based on a zone’s population relative to the region’s total population, as well as the zone's travel impedance to other zones relative to the travel impedance to all zones from that zone. The balancing factors ensure that each residential parcel (i.e., students in a home) are assigned a portion of the total school-seats in the region. This intuitive approached ensured the school-seats from each school is proportionally allocated to each zone based on the zone's relative population and travel impedance _and_ each value summed to the total number of opportunities in the region. Meaning, only a share of the school-seats at each school are allowed to be allocated. At some point during the reading course, I decided to continue onward with my academic journey and begin my PhD with Antonio Paez and as a Researcher with the Mobilizing Justice partnership in September 2021.

The competitive accessibility method developed, which preserves the zonal units, was introduced as “spatial availability" and published in _PLOS ONE_ as my first article of my PhD journey [@soukhovIntroducingSpatialAvailability2023]. At the time, we recognized that "spatial availability" is _singly-constrained_ from the perspective of opportunities, akin to the singly-constrained spatial interaction model in @wilson1971, in part thanks to Chris' comment to the effect of, "this looks a bit like Wilson's balancing factors". We adopted the term 'singly-constrained', without heavily scrutinizing the Wilson's framework (at the time). Instead I focused on the development of other skills: for instance, the publication of an open data paper demonstrating my growing knowledge of R and reproducible methods in _Environment and Planning B: Urban Analytics and City Science_ [@soukhovTTS2016RDataSet2023] then followed, then a multimodal extension of the spatial availability measure the following year [@soukhovMultimodalSpatialAvailability2024]. Along with years of thinking and reading about the equity assessment of transportation systems summarised in the (forthcoming) review of the literature @soukhovSearchingFairnessStandards2025 (and present in its long form [here](https://mobilizingjustice.ca/wp-content/uploads/2024/01/just-transportation-1.pdf)), the first work--that started back in that 2021 reading course--the article assessing the change in active (and motorized) school-seat accessibility after 10 years of school closures--was ultimately published in _Networks and Spatial Economics_ [@soukhovTenYearsSchool2025].

These efforts could have formed the basis of their own dissertation. However, I would have been unsatisfied--on account of the connections developed over the end of 2024 and throughout 2025. Through discussions with accessibility researchers at international and regional conferences over the course of the PhD program, as well as continued engagement with the literature, I found myself still searching for a satisfying understanding of how spatial interaction models and accessibility measures are connected. Was it only that the proportional allocation factor in @soukhovIntroducingSpatialAvailability2023, @soukhovMultimodalSpatialAvailability2024 and @soukhovTenYearsSchool2025 resemble Wilson's singly constrained balancing factor and otherwise our worlds are distinct, or was there something more?

As it turned out, spatial availability can be simplified to be mathematically equivalent to the singly constrained spatial interaction model. This means that one of the most widely used competitive accessibility measure--the 2SFCA--can also be understood as being proportion to Wilson's singly-constrained expression (NOTE: in @soukhovIntroducingSpatialAvailability2023, the mathematical equivalence between spatial availability per capita and 2SFCA is demonstrated). Following this logic, a total constrained version of accessibility, also fitting within Wilson's framework, can be established. The total constrained accessibility measure would require only information about the total number of opportunities in the region and at each destination, connecting travel costs between origins and destinations, and the opportunity being non-competitive in its definition. Conveniently, this case of the measure remains proportional to the conventional accessibility in Equation \ref{eq:access-unconstrained} but ensures the zonal accessibility values are proportional to the total in the region and maintains its units, i.e. the _number of accessible opportunities_ for each zone.

These connections came to be in discussions with Antonio Paez and Chris Higgins, along with Rafael H. M. Pereira who was visiting professor at the University of Toronto in the Fall of 2024 and Spring of 2025. From this position, a dissertation focused on spatial availability would be incomplete. So, this dissertation adopts a different flavour: the exploration of how spatial interaction models and accessibility measures can be connected, and how the introduction of units to accessibility measures can intuitively improve the understanding of accessibility analysis through a single empirical example.

This dissertation begins with this forthcoming (and submitted) work of @soukhovfamilyofaccessibility2025 which details the motivation and intuition of using the constraints from the spatial interaction modelling framework to establish measures of the "potential of opportunities" (or population) for spatial interaction. Both the singly constrained accessibility case (previously published in a different formulation as spatial availability) and, for the first time, the total constrained accessibility case is introduced and formulated.

In summary, this dissertation is based on or inspired by the following publications (listed in chronological order of their publication):

- **Soukhov, A.**, Pereira, R. H. M., Higgins, C. H. & Paez, A. (2025). A family of accessibility measures derived from spatial interaction principles. (Forthcoming - submitted to PLOS ONE).
- **Soukhov, A.**, Higgins, C. D., Páez, A., & Mohamed, M. (2025). Ten Years of School Closures and Consolidations in Hamilton, Canada and the Impact on Multimodal Accessibility. Networks and Spatial Economics. https://doi.org/10.1007/s11067-025-09677-z
- **Soukhov, A.**, Aitken, I. T., Palm, M., Farber, S., & Paez, A. (2025). Searching for fairness standards in the transportation literature. (Forthcoming - submitted to *Transportation*). An extended version, as a report, can be found here: https://mobilizingjustice.ca/wp-content/uploads/2024/01/just-transportation-1.pdf 
- **Soukhov, A.**, Tarriño-Ortiz, J., Soria-Lara, J. A., & Páez, A. (2024). Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes. PLOS ONE, 19(2), e0299077. https://doi.org/10.1371/journal.pone.0299077
- **Soukhov, A.**, & Páez, A. (2023). TTS2016R: A data set to study population and employment patterns from the 2016 Transportation Tomorrow Survey in the Greater Golden Horseshoe area, Ontario, Canada. Environment and Planning B: Urban Analytics and City Science, 23998083221146781. https://doi.org/10.1177/23998083221146781
- **Soukhov, A.**, Paez, A., Higgins, C. D., & Mohamed, M. (2023). Introducing spatial availability, a singly-constrained measure of competitive accessibility. PLOS ONE, 1–30. https:// doi.org/10.1371/journal.pone.0278468

In addition, several tangential projects undertaken alongside, but not directly related to, this dissertation have been published or are forthcoming, as listed below:

- **Soukhov, A.**, Foda, A., & Mohamed, M. (2022). Electric Mobility Emission Reduction Policies: A Multi-Objective Optimization Assessment Approach. Energies, 15(19), 6905. https://doi.org/10.3390/en15196905
- Leadbetter, I., Hernandez Cruz, A. E., Ravensbergen, L., & **Soukhov, A.** (2024). The Negative Impact of Parking Lots on Walkability. Findings, December. https://findingspress.org/article/127106-the-negative-impact-of-parking-lots-on-walkability
- Amorim Parga, J. P. F., **Soukhov, A.**, Arku, R. N., Higgins, C. D., & Páez, A. (2025). Democratic Access to our Cities: The impacts of recent changes to transit services in major Canadian metropolitan areas (Canada’s Urban Infrastructure Deficit: Toward Democracy and Equitable Prosperity,). University of Toronto’s School of Cities. https://stateofcitiessummit.ca/files/041224_Democratic-Access-to-our-Cities-The-Impacts-of-Recent-Changes-to-Transit-Services-in-Major-Canadian-Metropolitan-Areas.pdf
- **Soukhov, A.**, Mooney, N., & Ravensbergen, L. (2025). Exploring mobility of care with measures of accessibility. Journal of Transport Geography, 123(104050). https://doi.org/10.1016/j.jtrangeo.2024.104050
- Hernandez Cruz, A. E., Leadbetter, I., Ravensbergen, L., & **Soukhov, A.** (2025). Accessibility vs. Walkability: Exploring “Complete” Access to Care Destinations. (Forthcoming Journal Article).
- Bonsma-Fisher, M., **Soukhov, A.**, Rawal V., Chan, T. C. Y., & Saxe, S. (2025). When proximity to sustainable transportation infrastructure does not match accessibility to opportunities: A Canadian case study. (Forthcoming Book Chapter).

## The importance of accessibility; and conceptual issues interpreting 'unconstrained' access

Transportation is a special kind of land-use that compresses space and time, enabling connections, and hence potential interaction, between populations in different zones. But, as posed in @handy2020, "is accessibility an idea whose time has finally come?" This concept has been discussed in the literature for decades as the ulterior goal of transportation systems. Accessibility doesn't express mobility directly but instead captures the potential for meaningful mobility--the reachability of opportunities for people. Despite the utility of accessibility measures in both literature and practice to identify areas of high and low opportunity access, the interpretability of foundational accessibility measures, such as the Hansen-type accessibility measure [@hansen1959] (as in Equation \ref{eq:access-unconstrained}), can be challenging [@geursAccessibilityEvaluationLanduse2004; @millerAccessibilityMeasurementApplication2018; @santanapalacios2022].

By definition, accessibility is the product of the sum of opportunities that can be reached through some travel impedance, meaning each zonal value is sensitive to both inputs in a way that isn't constrained by any known system. For example, consider comparing the accessibility values for a region between two years where the number of opportunities remained constant but the travel behaviour dramatically changed. In this case, both years would have travel impedance functions of different forms. As a result, the zone values would have different units, one in opportunities-weighted-travel-impedance-value-from-form-1 and another in units of opportunities-weighted-travel-impedance-value-from-form-2. These values would exist within different ranges, and while they could be normalized and compared, this process is an additional treatment needed to locate meaning.

Also consider if interested in comparing one region to another region. The number of opportunities varies between the two regions (e.g., it may be expected that larger regions have more opportunities than smaller ones). For this reason, raw accessibility scores calculated by conventional 'unconstrained' methods are not necessarily comparable without some form of normalization. It will likely be that the central areas in a more urban area will have larger opportunities-weighted-travel-impedance-values than central area of a less urban region. Again, to facilitate comparision between the regions, normalisation is needed which is an additional treatment needed to locate meaning. As with many spatial issues, these problems are compounded by the Modifiable Areal Unit Problem (MAUP) and Modifiable Temporal Unit Problem (MTUP), where the zonal and temporal units may differ between comparison periods, further obscuring results.

Aside from temporal and regional comparisons, gleaning interpretation of raw 'opportunities-weighted-travel-impedance-value' scores within the same temporal and spatial region is also challenging. To move closer to an intuitive interpretation, a balancing factor that is sensitive to the known information about the transportation system can be introduced. This would allow the zonal value of 'how many opportunities can be accessed' to be more plainly contemplated--as the units of travel impedance are set aside. This would allow linking this value that has been unified to its accessibility 'units' directly to something meaningful (e.g., the number of opportunities needed for a specific outcome). In this way, accessibility values that have been reunited with their units become more straightforward and interpretable without any treatment required. This form of intuitive interpretation of the units is the central aspect of this dissertation.

## Aims

To reiterate, the primary contribution of this dissertation is demonstrating how the units of accessibility--the number of opportunities that can potentially be reached--can be preserved within the zonal values of the accessibility measure. This contribution offers a practical way to enhance accessibility as a transportation and land-use planning tool, making it more interpretable and potentially more useful for decision-making.

This dissertation offers the four following contributions:

First, it defines accessibility as the _potential_ for spatial interaction by explicitly linking it the spatial interaction literature, and providing a review of how the two streams of literature diverged. Next, it explores how spatial interaction modelling literature maintains units in its outputs by introducing system and zonal constraints through balancing factors, and demonstrates how these constraints can be applied to the accessibility context.

Second, the total constraint is introduced in the accessibility context. The resulting total constrained accessibility measure is conceptually equivalent in magnitude to the Hansen-type accessibility measure (Equation \ref{eq:access-unconstrained}) but is expressed in units of accessible opportunities or accessible population. This approach can help ensure that accessibility measures remain consistent and comparable across different areas and time periods, addressing the issue of unit discrepancies that arise in accessibility analyses.

Third, the single constraint is introduced, which accounts for competition among either the populations for opportunities or the opportunities supplied for the population, as well as ensuring the total constraint is met too. The singly constrained accessibility measure can be understood as a competitive measure and its per capita variant is mathematically equivalent to the Two Step Floating Catchment Area (2SFCA) method [@luo2003]. In this way, singly constrained accessibility can offer a more accurate understanding of localized competition for opportunities (or for population spatially interacting with opportunities).

Fourth, the multi-modal extension of both total- and singly- constrained accessibility is specified. Specifically, these extensions demonstrate how these measures can be adapted to consider different groups with different travel impedance functions (i.e., be it multiple modes of transport or populations with different travel likelihoods) accessing the same opportunities. This multi-modal extension can be used to interpret the accessibility gap between groups, illustrating how many more opportunities can be reached by one group compared to another.

In sum, the intuition, methodology and empirical examples offered by this dissertation aims to increase the interpretability of resulting values from accessibility measures. This opens the doors for accessibility analysis that is, hopefully, easier to communicate and tie to practical outcomes and goals. The first aim is addressed in Chapter 1, and the second, third and fourth aims are formulated in Chapter 2, outlined again but with respect to an empirical example of parkland area in Toronto in Chapter 3, and expressed as in unimodal accessibility to parkland area and population in Chapter 4, and multimodal accessibility to parkland area and population in Chapter 5.

## A practical case study: Toronto and publicly owned and opertated parkland

To demonstrate and compare the constrained accessibility measures presented in the Chapter 1 and 2, an empirical example is developed. 

As mentioned: an advantage of the _constrained_ accessibility measures is in the resulting values having tangible units. Static metrics of opportunity supply for a spatial zone are often used in planning, e.g. the number of jobs located in a census block or the number of hospital beds per person in a neighbourhood. Instead, constrained accessibility measures yield a value of the number of _accessible_ opportunities per zone or per capita (in a zone). In effect, resulting values can be interpreted using the inuition of opportunity supply metrics but also fold in transportation-land-use-interaction assumptions e.g., how the population may interact with that opportunity.

To showcase this practical advantage: a case study reporting the accessibility to green space that is owned and operated by the City of Toronto is detailed. As specified in the City of Toronto's most recent parkland strategy report, green space is vital for wellbeing as it promotes physical health by encouraging outdoor activity, enhances mental health through reduction in anxiety and stress, and fosters social connection by providing communical space for all to use [@toronto_parkland2019]. In this strategy, the City also defines priority areas in need of additional parkland area supply: namely areas of lower income and below an accessible supply of 12 sqm per capita.

From this perspective, the city has identified greenspace as a crucial public good and benchmarking its accessible supply (i.e., constrained accessibility), is an important step towards the identification of priority areas. From this perspective, accessibility of parkland area from areas in the city, and accessibility of population to parks, is a fitting practical case study for the empirical exploration of the constrained accessibility measures.

Furthermore, greenspace is an interesting opportunity as it can be considered from both a non-competitive and competitive vantage points. From a non-competitive perspective, one would assume that parkland is essentially inexhaustible, meaning my neighbour and I can both enjoy the same amount of parkland area without reducing each other’s accessibility. The total constrained accessibility measure would yield values that reflect the share of _total_ parkland area that you can reach, and that would be similar to the shares allocated to all your neighbours that can reach the same parks. Conversely, from a competitive perspective, park space can be seen as exhaustible or finite--if a park becomes overcrowded, my ability to use it diminishes as more people share the same area. This intuition is captured in the singly-constrained accessibility measure, which will yield a value that reflects the share of parkland area only from _parks that are reachable_ as determined by population demand for these parks (as well as travel cost). 

In sum, to demonstrate how constrained accessibility may work in practice, this dissertation compares the introduced constrained accessibility measures to the traditional (unconstrained) measure in the context of parkland accessibility in Toronto. By highlighting the interpretative advantage of the constrained measure, the dissertation aims to use the empirical example to demonstrate how values of constrained accessibility can be translated to supply thresholds and be potentially used by policymakers to improve transportation and land-use. 

## Overview of methods 

### Reproducible statement

This dissertation adopts principles of open and reproducible research practices. For instance, all analysis conducted and write-up create was done using using R and RStudio, with the complete workflow--including code, data, and text of the final dissertation manuscript, publicly accessible in the the accompanying [repository](https://github.com/soukhova/Dissertation). To ensure full transparency and assist in replicability, the repository features a landing page with reproducibility guidelines and detailed documentation within the data-raw directory. 

Adopting these principles not only underscores the importance of reproducibility in scientific inquiry, as discussed in [@brunsdon2021opening; @paez2021open; @arribas2021open], but also provides a tangible framework for reuse by facilitating others to build on these efforts in the academic realm and by facilitating the translation of methods and findings into practice.
	
## Chapters outline

As mentioned, this dissertation is divided into six chapters. Chapter 1 provides the conceptual and historical framework of this dissertation as written in the first half of the manuscript titled _"Family of accessibility measures derived from spatial interaction principles"_ submitted to _PLOS ONE_. This chapter outlines the evolution of the spatial interaction literature, how accessibility literature diverged, and introduces the total and single constraints using numeric examples. 

Chapter 2 outlines the family of accessibility measures (unconstrained, total constrained, singly constrained, and doubly constrained accessibility) using a simple toy example. It is the second part of the forthcoming manuscript "Family of accessibility measures derived from spatial interaction principles"_ submitted to _PLOS ONE_. As well as additional writing detailing the context for and the formulation of a multimodal extension of each measure, inspired writing from the published _"Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes"_ paper is also included [@soukhovMultimodalSpatialAvailability2024], as well as inspiration taken from the the published paper focusing on a singly mode _"Introducing spatial availability, a singly-constrained measure of competitive accessibility"_ [@soukhovIntroducingSpatialAvailability2023].

After the first two chapters, all work in this dissertation is original. 

Chapter 3 details the empirical example of parkland accessibility in Toronto developed in chapters four and five of this dissertation. Namely the origins, the different zonal levels of aggregation, qualities about the destinations, and the routing and trip length assumptions are detailed. The unconstrained and constrained accessibility measures used in the chapters four and five are briefly summarised in context with the empirical example. 

Chapter 4 calculates a unimodal example of parkland accessibility and accessible population (to parkland area) for the city of Toronto, using the unconstrained, total and singly constrained accessibility. Results using different measures are compared, discussed and contextualized. Discussion of the supply ratios is inspired by the paper focused on unimodal, namely the empirical example in is inspired by the published paper on the topic of multimodal competitive accessibility measures, namely  _"Introducing spatial availability, a singly-constrained measure of competitive accessibility"_ [@soukhovIntroducingSpatialAvailability2023] 

Chapter 5 builds on the preceding chapter, and extends the example to consider multiple modes: walking, as well as cycling, transit and car. Multimodal accessibility to parkland (and to population from parkland) is detailed. Results are compared to the unimodal example, and compared, discussed and contextualised. Discussion of competitive gains is inspired by the published paper on the topic of multimodal competitive accessibility measures, namely  _"Ten Years of School Closures and Consolidations in Hamilton, Canada and the Impact on Multimodal Accessibility"_ [@soukhovTenYearsSchool2025] and _"Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes"_ [@soukhovMultimodalSpatialAvailability2024]

Chapter 6 concludes the work by summarising each chapter and outlining directions for future research.

<!--

BEGINNING HERE ADD THE DOCUMENTS THAT CONFORM THE MAIN BODY OF THE THESIS

-->

```{r eval=!knitr::is_latex_output(), child=here::here("01-chap1.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("02-chap2.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("03-chap3.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("04-chap4.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("05-chap5.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("06-conclusion.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("99-references.Rmd")}

```
