---
title: 'Reuniting Accessibility Measures with Spatial Interaction Principles'
author: 'Anastasia Soukhov'
date: 'August 2025'
university: 'McMaster University'
faculty: 'Faculty of Science'
department: 'School of Earth, Environment and Society'
subject: 'Transport Geography'
supervisor: 'Antonio Páez'
bdegree: 'B.Eng.' # Bachellor's Degree
mdegree: 'M.A.Sc.'   # Master's Degree (if any)
# This is the degree you are aiming for with this thesis
degree: 'Doctor of Philosophy'
knit: bookdown::render_book
site: bookdown::bookdown_site
output:
  macdown::thesis_pdf: default 
  keep_tex: true
lay_abstract: |
  One of the primary goals of transportation systems is to connect people with opportunities--such as jobs, essential services and community. However, traditional transportation planning practice has largely emphasized mobility, focusing on metrics like distance traveled or travel time, rather than accessibility (the number of opportunities people can reach). Although researchers have long advocated for a shift from mobility-based approaches to accessibility-focused ones, this transition has yet to be fully realized due to several persistent challenges. A quiet barrier is methodological: the absence of consistent and tangible units for measuring accessibility. This thesis seeks to address this barrier by examining how accessibility methods relate to conventional mobility-based techniques, and by exploring how constraints--adapted from spatial interaction models--can be used to return meaningful units to accessibility values and their analysis. This is done by reviewing the related spatial interaction modeling and accessibility literature, outlining a family of 'constrained' accessibility measures, and detailing an empirical example of accessible parkland area and population in the City of Toronto across all measures.
  
abstract: |
  Transportation systems play a fundamental role in facilitating access between people and opportunities across place. Place-based accessibility can be defined as the potential of opportunities for spatial interaction. For decades, transportation planning research and practice have relied on mobility-based metrics, such as kilometers traveled, emissions released, or vehicle counts, measures that reflect realised movement. In contrast, accessibility focuses on potential movement, such as the number of opportunities that can be reached within given constraints. Although there has been growing interest in shifting from mobility-based methods to access-focused planning approaches, a few barriers remain. One significant issue is methodological, namely, the lack of clarity in how accessibility values are interpreted. There are several approaches to match values with meaning, but this dissertation proposes a preceding step: clarifying the units of accessibility by introducing proportionality constants derived from spatial interaction principles.

  In response, this monographic-style dissertation is organized into six chapters. In the first chapter, a review of how accessibility literature largely diverged from the spatial interaction literature is detailed. This lays the foundation for how the addition of proportionality constants that return the units to the measure and balance them to reflect known constraints in the system may be useful. In the second chapter, using a synthetic example, the total constrained accessibility measure is introduced. This measure is formally proportional to unconstrained accessibility (i.e., Hansen-type accessibility measure commonly used in literature) but with results consistently expressed in units of accessible opportunities or units of accessible population. Then, using the same synthetic example, the single constrained accessibility measure is introduced. This measure considers competition for opportunities and could be understood as related to the popular measures of competitive accessibility. Multimodal extensions as well as the expression of these measure in terms of ‘market potential'--the accessible population at a zone--is also introduced. 
  
  In the third chapter, motivation, data and methods of an empirical example of the parkland area and population in the City of Toronto are detailed. All members of the family of accessibility measures introduced in this work are compared and contrasted in the subsequent chapters using this empirical example. In the fourth chapter, a unimodal walking-only case considering a binary impedance function and all population is presented, where the accessible parkland area and accessible population in Toronto is quantified. Differences in the unconstrained (i.e., conventional method) and constrained measures are detailed. Fifth, a multimodal case of accessible parkland area and accessible population is detailed, considering walking, cycling, transit, and car modes. Similarly, differences in the unconstrained and constrained measures are detailed. In the sixth and final chapter, a summary of key points across all chapters is provided along with a thorough discussion on how constrained accessibility improves communicability and details on future lines of potential investigation.
  
# Note that a tab is needed on the line after the `|`. Subsequent paragraphs need to be indented too.
acknowledgements: |
  Oh the places this dissertation has taken me.
  
  First and foremost, I want to express gratitude to my PhD supervisor Antonio Páez, for your unwavering guidance throughout this journey. From co-working coding & writing sessions, bi/weekly check-ins and countless opportunities to present my work in regional and international venues: the freedom and support you've provided to allow me to develop ideas, my academic voice, and establish myself over the past four years have made the completion of this dissertation possible. I also want to sincerly thank Moataz Mohamed, my MASc supervisor, who first folded me into the academic world. I want to also thank Christopher D. Higgins, who worked alongside Antonio, Moataz and myself on one of the earliest projects of my PhD journey--work that ultimately set the stage for this dissertation. Thank you!

  I also want to express gratitude for the Mobilizing Justice Partnership, led by Steven Farber and supported academically in its early years by Ignacio Tiznado-Aitken and Matthew Palm--thank you for the guidance and support in framing my understanding of transportation equity and justice. I'm also especially grateful for the Partnerships'connection with the University of Toronto and the opportunity to work with others with similar research passions; Madeleine Bonsma-Fisher, João Pedro Figueira Amorim Parga, and Robert Nutifafa Arku. Also, a major thank you to the Mobilizing Justice Partnership's network of transportation practitioners committed to advancing equity. Through this network, I had the opportunity to engage with the City of Toronto's work in developing the Transportation Equity Policy Framework, led by Lorina Hoxha. As a member of the technical advisory committee, I witnessed firsthand the dedication of public servants pushing the boundaries of defining equity analysis in practice. I was grateful to contribute feedback on methods in various forums, faciliate workshops and participate in some lively discussions with fellow accessibility fanatics--Michael Hain, Bryce Sherman, and Herman Hui. These conversations inspired me to apply the methods introduced in this dissertation to a practical example they were exploring. I extend my sincere gratitude to all of them.

  I would also like to extend my gratitude to those at McMaster University. I am particularly grateful to Léa Ravensbergen, whose mentorship opened the door to my first (and subsequent) co-supervisions of passionate undergraduate students at McMaster University: Nicholas Mooney, Isla Leadbetter, and Angel Emmanuel Hernandez Cruz. I also want to thank the growing community of transportation reseachers--the Friends of "Transport Fest"--Darren Scott and his students Yehui Yin, Haoran Xu, and Sadia Chowdhury, along with my academic siblings Bruno Dos Santos and Amanda Rorat and visiting PhD students Javier Tarriño-Ortiz and Carlos Marigil Alba, along with the comraderie of Léa students: Alexandra Sbrocchi, Dastar Bhoday, Domenica De Iuliis, and Moataz's students particuarly: Sadia Tasnim, Ali Shehabeldeen and Ahmed Foda. Though I completed the PhD program and this dissertation while living in Toronto, you all made visiting campus that much more worthwhile. Your perspectives sharpened my thinking and helped me see both familiar and new problems in different ways.

  To my international friends and collaborators, a big thank you! I'm especially grateful to Rafael H. M. Pereira who was looped into my work during his stay as an visiting professor at UofT in Fall 2024 / Spring 2025. I'm also grateful to everyone who made my two spring research stays at TRANSyT at the Universidad Politécnica de Madrid in 2023 and 2024 productive, tremendously fun, and ultimately unforgettable. Thank you to Julio A. Soria-Lara, Alberto Rojas-Rivero, Manuel Pérez, Raúl F. Elizondo-Candanedo, Miguel L. Navarro-Ligero, Amor Ariza-Álvarez, and all the other lab mates at TRANSyT for your support, collaboration, and friendship. And to the friends, colleagues, and co-conspirators I've had the good fortunte of connecting with, gaining inspiration from, and recieving feedback from at international conferences (i.e., NARSC, AAG, TRB, WCTRS, and NECTAR): thank you. To the co-organizers of the Transportation Justice group at AAG and the members of the NECTAR Board--thank you for creating spaces that foster both a passion for advancing transportation research and a sense of academic community. These international connections made the pursuit of a PhD all the more worthwhile.

  To my partner in life (and now husband), Aidan: thank you for your patience, love, and unwavering belief in me and in the goodness in the pursuit of knowledge for a better world. From encouraging me to take the plunge into the PhD program, to supporting my obsessive late-night working sessions before (often self-imposed) deadlines, to planning our wedding while I was furiously writing away on this dissertation, to standing by through weeks and months apart during conferences and research stays--your support has been constant. Your belief in me, and your belief in the value of continuing this academic journey, means more than I can ever fully express.

  This work stands on the shoulders of many: many more than I can name possibly list. Thank you to all before me whose work helped create the path I was lucky enough to find myself on. Thank you for making it possible for me to inch just a bit further, and to contribute something novel to our field.

declarations: |
 I am the primary author of this dissertation. I also authorize McMaster University to lend this dissertations to other institutions or individuals for the purpose of scholarly research as well as reproduce it in total or in part for the purpose of scholarly research.
  
# quotation: 'There is no such thing as a new idea. It is impossible. We simply take a lot of old ideas and put them into a sort of mental kaleidoscope... they make new and curious combinations...'
# 
# # Author of the quotation
# qauthor: 'Mark Twain'
# Specify the location where the bibliography file is found below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include = FALSE}
# This chunk ensures that the macdown package is
# installed and loaded. This macdown package includes
# the template files for the thesis.
if(!require(remotes))
  install.packages("remotes", repos = "http://cran.rstudio.com")
if(!require(macdown))
  remotes::install_github("paezha/macdown")
library(macdown)
```

<!--
The acknowledgements, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
-->

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Overview {.unnumbered}

This work is a monograph-style dissertation comprising six chapters that together explore the development and application of a family of accessibility measures derived from the principles of spatial interaction to generate constrained accessibility values. These measures are applied to the empirical case study of parkland area and population in the City of Toronto and discussed. In sum: this dissertation offers theoretical and methodological advancements to the field of accessibility, along with practical empirical analysis to explore how these measures may be interpreted in practice.

The purpose of this introductory chapter is to provide an overview of the structure and logic to the organisation of the dissertation. It outlines the conceptual issue of accessibility measures being interpreted as indicators, the dissertation's development trajectory, the aims of the dissertation, overview of the empirical data, reproducibility statement, and a detailed outline of the chapters. 

## The importance of accessibility; and conceptual issues interpreting 'unconstrained' access

Transportation is a special kind of land-use that compresses space and time, enabling connections, and hence potential interaction, between populations in different zones. But, as posed in @handy2020: _"is accessibility an idea whose time has finally come?"_. 

Accessibility is a concept has been discussed in the literature for decades as the ulterior goal of transportation systems. Accessibility doesn't express mobility directly but instead captures the potential for meaningful mobility: the reachability of opportunities for people. Despite the utility of accessibility measures in both literature and practice to identify areas of high and low opportunity access, the interpretability of foundational accessibility measures, such as the Hansen-type accessibility measure [@hansen1959], can be challenging [@geursAccessibilityEvaluationLanduse2004; @millerAccessibilityMeasurementApplication2018; @santanapalacios2022].

By definition, accessibility is the product of the sum of opportunities that can be reached through some travel impedance, meaning each zonal value is sensitive to both inputs in a way that is not constrained by any known system. For example, consider comparing the accessibility values for a region between two years where the number of opportunities remained constant but the travel behaviour dramatically changed. In this case, both years would have travel impedance functions of different forms. As a result, the zone values would have different units, one in _opportunities-weighted-travel-impedance-value-from-form-1_ and another in units of _opportunities-weighted-travel-impedance-value-from-form-2_. These values would exist within different ranges, and while they could be normalized and compared, this process is an additional treatment needed to locate meaning.

Similarly, comparisons between regions, especially those of different sizes or opportunity distributions, can be misleading. For instance, even if using the same impedance functions and zones of analysis are selected, areas with more opportunities will have higher accessibility scores than areas with fewer opportunities even if the relative 'ease' of access is similar. For this reason, raw accessibility scores calculated by conventional methods are not necessarily comparable without some form of normalization.

To address these interpretability challenges, this dissertation proposes a constrained formulation of accessibility that restores meaningful units to the measure. By introducing a proportional allocation mechanism derived from spatial interaction principles, accessibility values can be made directly interpretable. Each zonal value then reflects a bounded share of total opportunities (e.g., school-seats, jobs, hectares of parkland) or total population, grounded in tangible units. This clearer basis for interpretation can potentially improve the clarity in equity evaluations, identifying spatial mismatches, and designing interventions, among other practical applications.

This dissertation is centrally concerned with this issue: how to reconnect accessibility measures to interpretable, bounded, and theoretically consistent units using the framework of spatial interaction modeling.

## A note on this dissertation's evolution

The motivation to address the issue of accessibility interpretability did not emerge in the abstract, but rather through a series of practical encounters. This research evolved from early attempts to make sense of real-world planning challenges--particularly those involving equity in accessibility to school-seats. In this way, this monographic-style dissertation has come together non-chronologically. 

Its origins begin in a reading course with Dr. Antonio Páez as a Masters of Applied Science (MASc.) student under the supervision of Dr. Moataz Mohammed. In this reading course, I was introduced to a dataset on public school closures in Hamilton. The closures raised questions about how students' access to educational opportunities might change, particularly for those within walking distance of the affected schools. This led to an investigation into the spatial distribution of school-seat opportunities and how their availability could be measured using place-based accessibility indicators. 

Conventional accessibility measure are typically represented as the sum of the number of opportunities $O_j$ weighted by the travel impedance function $f(c_{ij})$ of some travel cost $c_{ij}$ for a set of zones of origin $i$ $i...I$ and destination $j$ $j...J$. Equation \ref{eq:access-unconstrained} represents this general form:

```{=tex}
\begin{equation}
\label{eq:access-unconstrained}
S_i = \sum_i O_j f(c_{ij}) 
\end{equation} 
```

While widely used, such formulations often produce values that are effectively interpreted without units, making them difficult to tie to real-world quantities. Motivated by the question, _how did the number of potentially reachable school-seats change for households across space?_, I saught a more interpretable measure that retained physical units (e.g., number of school-seats) that still behaved intuitively like a conventional accessibility measure. 

To address these challenges, a new method was developed in collaboration with co-authors Chris Higgins, Moataz Mohamed and Antonio Páez. The approach, ultimately named "spatial availability" uses proportional allocation factors to distribute opportunities (school-seats) based on both the relative population of origin zones and their travel impedance values. These proportional allocation factors ensure that each residential parcel receives a proportional share of the regional school-seat total, preserving both the total number of school-seats in the region and the number of school-seats at each school as well as the logic of accessibility. This formulation was published in _PLOS ONE_ [@soukhovIntroducingSpatialAvailability2023]. Its multimodal extension followed in the next year [@soukhovMultimodalSpatialAvailability2024], and the school-seat case study itself, was published in _Networks and Spatial Economics_ investigating how active (and motorized) spatial availability to school-seats changed over a decade of school closures in Hamilton [@soukhovTenYearsSchool2025].

While these individual contributions could have formed the basis for a manuscript-based dissertation on "spatial availability", they did not fully resolve the core theoretical questions that emerged. Namely, the "spatial availability" approach was presented to transportation researchers at international conferences and seminars, and a recurring curiosity remained: _To what extent can accessibility measures be understood within the spatial interaction modeling framework?_

Initially, we assumed that spatial availability's constraining mechanism only resembled Wilson's balancing factors (in part thanks to Chris' early comment to the effect of _"this looks a bit like Wilson's balancing factors"_). However, deeper theoretical investigation revealed that spatial availability is mathematically equivalent to the singly-constrained spatial interaction model, though of course, derived from different paths. Further, this also meant that the most widely used competitive accessibility measure--the Two Step Floating Catchment Area (2SFCA)--can be understood as being linked to Wilson's singly-constrained expression (NOTE: in @soukhovIntroducingSpatialAvailability2023, the mathematical equivalence between spatial availability per capita and 2SFCA is demonstrated). Following this logic, a total constrained version of accessibility, also fitting within Wilson's framework and proportional to the conventional accessibility in Equation \ref{eq:access-unconstrained}, could be developed along with multimodal extensions and 'market potential' variants.

These theoretical connections to spatial interaction warranted a more comprehensive and unified exploration. As a result, this dissertation is monographic in style, and explores how spatial interaction models and accessibility measures can be connected more explicitly. It uses an original empirical case study across all cases and variants of measures to compare, contrast and comment on how this framework improves accessibility's interpretability and policy relevance while drawing linkages between existing measures--such as Equation \ref{eq:access-unconstrained} and the 2SFCA approach--unifying popular non-competitive and competitive approaches.

## Aims

The primary aim of this dissertation is to develop accessibility measures that preserve its units of analysis (i.e., the number of opportunities that can potentially be reached) within zonal outputs by being grounded in spatial interaction principles. By retaining meaningful units and enabling probabilistic interpretation based on known constraints, this work offers a theoretically rigourous framework to measure and interpret accessibility values, enhancing clarity in interpretation relative to conventional measures and, inuitive usefulness for decision-making.

This dissertation makes the four following contributions:

First, it defines accessibility as the _potential_ for spatial interaction by explicitly linking it to the spatial interaction literature. It reviews early spatial interaction research which advances the Newtonian gravitation analogy ultimately applied in @hansen1959's work, as well as @wilson1971's breakthrough entropy-maximisation model building approach, to examine how the accessibility and spatial interaction literatures later diverged. 
However, spatial interaction modelling offers valuable insights for accessibility research: most notably, the ability to preserve interpretable units through the introduction of system-wide and zonal constraints, operationalised by balancing factors. This insight sets the theoretical ground for the reformulation of spatial interaction model's 'balancing factors' as 'proportional allocation' factors in the case of accessibility. 

Second, the total constraint--understood as the total flow constraint from spatial interaction modeling--is introduced in the accessibility context. The resulting total constrained accessibility measure is conceptually equivalent in magnitude to the Hansen-type accessibility measure (Equation \ref{eq:access-unconstrained}) but is expressed in units of accessible opportunities or accessible population. This approach intuitively ensures the resulting values remain consistent and comparable across different scenarios (pending appropriate system knowns), addressing the quiet issue of unit discrepancies that arise in accessibility analyses.

Third, the single constraint is introduced, which accounts for competition among either the population for opportunities or the opportunities supplied for the population. It also ensures the total constraint is maintained. The singly constrained accessibility measure can be understood as a competitive measure and its per capita variant is mathematically equivalent to the 2SFCA method [@shen1998;@luo2003]. In this way, singly constrained accessibility can offer a more accurate understanding of localized competition for opportunities (or for population potentially spatially interacting with opportunities).

Fourth, the multimodal extension of both total- and singly- constrained accessibility is specified. Specifically, these extensions demonstrate how these measures can be adapted to consider different groups with different travel impedance functions (i.e., be it multiple modes of transport or populations with different travel likelihoods) accessing the same opportunities. This multi-modal extension can be used to interpret the accessibility gap between groups with different modalities, illustrating how many more opportunities can be reached by one group compared to another.

The first aim is addressed in Chapter 1, and the second, third and fourth aims are formulated in Chapter 2, outlined again but with respect to an empirical example of parkland area in Toronto in Chapter 3, and expressed as in unimodal accessibility to parkland area and population in Chapter 4, and multimodal accessibility to parkland area and population in Chapter 5.

Together, the intuition, methodology and empirical examples offered by this dissertation aims to increase the interpretability of resulting values from accessibility measures. This work opens the doors for accessibility analysis that is, hopefully, easier to communicate and tie to practical outcomes and goals. 

## A practical case study: Toronto and publicly owned and operated parkland

To demonstrate and compare the constrained accessibility measures formalised in Chapter 1 and 2, an empirical example is developed. 

As mentioned: an advantage of the _constrained_ accessibility measures is in the resulting values having tangible units. Static metrics of opportunity supply for a spatial zone are often used in planning, e.g. the number of jobs located in a census block or the number of hospital beds per person in a neighbourhood. Instead, constrained accessibility measures yield a value of the number of _accessible_ opportunities per zone or per capita (in a zone). In effect, resulting values can be interpreted using the intuition of opportunity supply metrics but also fold in transportation-land-use-interaction assumptions e.g., how the population may interact with that opportunity.

To showcase this practical advantage: a case study reporting the accessibility to green space that is owned and operated by the City of Toronto is detailed. As specified in the City of Toronto's most recent parkland strategy report, green space is vital for wellbeing as it promotes physical health by encouraging outdoor activity, enhances mental health through reduction in anxiety and stress, and fosters social connection by providing common space for all to use [@toronto_parkland2019]. In this strategy, the City also defines priority areas in need of additional parkland area supply: namely areas of lower income and below an accessible supply of 12 sqm per capita.

From this perspective, the city has identified greenspace as a crucial public good and benchmarking its accessible supply (i.e., constrained accessibility), is an important step towards the identification of priority areas. From this perspective, accessibility of parkland area from areas in the city, and accessibility of population to parks, is a fitting practical case study for the empirical exploration of the constrained accessibility measures.

Furthermore, greenspace is an interesting opportunity as it can be considered from both a non-competitive and competitive vantage points. From a non-competitive perspective, one would assume that parkland is essentially inexhaustible, meaning a person and their neighbours can enjoy the same amount of parkland area without reducing each other's accessibility. The total constrained accessibility measure would yield values that reflect the share of _total_ parkland area that you can reach, and that would be similar to the shares allocated to all your neighbours that can reach the same parks. Conversely, from a competitive perspective, park space can be seen as exhaustible or finite, i.e., if a park becomes overcrowded, one's ability to use it diminishes as more people share the same area. This intuition is captured in the singly-constrained accessibility measure, which will yield a value that reflects the share of parkland area only from _parks that are reachable_ as determined by population demand for these parks (as well as travel cost). 

To demonstrate how constrained accessibility may work in practice, this dissertation compares the introduced constrained accessibility measures to the conventional (unconstrained) measure in the context of parkland accessibility in Toronto. By highlighting the interpretative advantage of the constrained measure, the dissertation aims to use the empirical example to demonstrate how values of constrained accessibility can be translated to supply thresholds and be potentially used by policymakers to improve transportation and land-use. 

## Reproducible statement

This dissertation adopts principles of open and reproducible research practices. For instance, all analysis conducted and write-up was done using using R and RStudio, with the complete workflow; including code, data, and text of the final dissertation manuscript, publicly accessible in the the accompanying [repository](https://github.com/soukhova/Dissertation). To ensure full transparency and assist in replicability, the repository features a landing page with reproducibility guidelines and documentation within the data-raw directory. 

Adopting these principles not only underscores the importance of reproducibility in scientific inquiry, as discussed in [@brunsdon2021opening; @paez2021open; @arribas2021open], but also provides a tangible framework for reuse by facilitating others to build on these efforts in the academic and applied realm.
	
## Chapters outline

This monograph-style dissertation is divided into six chapters. 

Chapter 1 provides the conceptual and historical framework of this dissertation as written in the first half of the manuscript titled _"Family of accessibility measures derived from spatial interaction principles"_ submitted to _PLOS ONE_. This chapter outlines the evolution of the spatial interaction literature, how accessibility literature diverged, and introduces the total and single constraints using numeric examples. 

Chapter 2 outlines the family of accessibility measures (unconstrained, total constrained, singly constrained, and doubly constrained accessibility) using a simple toy example. It is the second part of the forthcoming manuscript _"Family of accessibility measures derived from spatial interaction principles"_ submitted to _PLOS ONE_. As well as additional writing detailing the context for and the formulation of a multimodal extension of each measure, inspired writing from the published _"Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes"_ paper is also included [@soukhovMultimodalSpatialAvailability2024], as well as inspiration taken from the the published paper focusing on a singly mode _"Introducing spatial availability, a singly-constrained measure of competitive accessibility"_ [@soukhovIntroducingSpatialAvailability2023].

After the first two chapters, all work in this dissertation is original. 
Chapter 3 details the empirical example of parkland accessibility in Toronto developed in chapters 4 and 5 of this dissertation. Namely the origins, the different zonal levels of aggregation, qualities about the destinations, and the routing and trip length assumptions are detailed. The unconstrained and constrained accessibility measures used in the chapters four and five are briefly summarised in context with the empirical example. 

Chapter 4 calculates a unimodal example of parkland accessibility and accessible population (to parkland area) for the city of Toronto, using the unconstrained, total and singly constrained accessibility. Results using different measures are compared, discussed and contextualized. Discussion of the supply ratios is inspired by the paper focused on unimodal, namely the empirical example in is inspired by the published paper on the topic of multimodal competitive accessibility measures, namely  _"Introducing spatial availability, a singly-constrained measure of competitive accessibility"_ [@soukhovIntroducingSpatialAvailability2023] 

Chapter 5 builds on the preceding chapter, and extends the example to consider multiple modes: walking, as well as cycling, transit and car. Multimodal accessibility to parkland (and to population from parkland) is detailed. Results are compared to the unimodal example, and compared, discussed and contextualised. Discussion of competitive gains is inspired by the published paper on the topic of multimodal competitive accessibility measures, namely  _"Ten Years of School Closures and Consolidations in Hamilton, Canada and the Impact on Multimodal Accessibility"_ [@soukhovTenYearsSchool2025] and _"Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes"_ [@soukhovMultimodalSpatialAvailability2024]

Chapter 6 concludes the work by summarising each chapter and outlining directions for future research.

<!--

BEGINNING HERE ADD THE DOCUMENTS THAT CONFORM THE MAIN BODY OF THE THESIS

-->

```{r eval=!knitr::is_latex_output(), child=here::here("01-chap1.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("02-chap2.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("03-chap3.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("04-chap4.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("05-chap5.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("06-conclusion.Rmd")}

```
<!--
```{r eval=!knitr::is_latex_output(), child=here::here("99-references.Rmd")}

```
-->