---
title: 'A family of accessibility measures: bringing practical interpretation to accessibility analysis'
author: 'Anastasia Soukhov'
date: 'June 2025'
university: 'McMaster University'
faculty: 'Faculty of Science'
department: 'School of Earth, Environment and Society'
subject: 'Geography'
supervisor: 'Antonio Paez'
bdegree: 'B.Eng.' # Bachellor's Degree
mdegree: 'M.A.Sc.'   # Master's Degree (if any)
# This is the degree you are aiming for with this thesis
degree: 'Doctor of Philosophy'
knit: bookdown::render_book
site: bookdown::bookdown_site
output:
  macdown::thesis_pdf: default 
  keep_tex: true
lay_abstract: |
  One of the primary goals of transportation systems is to connect people with opportunities such as jobs and essential services. However, traditional transportation planning has largely emphasized mobility, focusing on metrics like distance traveled or travel time, rather than accessibility (the number of opportunities people can reach). While researchers in the field have long advocated for a shift from mobility-based approaches to accessibility-focused ones, this transition has yet to be fully realized due to several challenges. A key barrier is methodological: the absence of standardized units for measuring accessibility. This thesis seeks to address that gap by examining how accessibility methods relate to conventional mobility-based techniques and by exploring how the concept of "constraints" can be adapted from these techniques to (re)introduce measurable units into accessibility assessments.
  
abstract: |
  Transportation systems play a fundamental role in facilitating access between people and opportunities across place. Spatial place-based accessibility, otherwise known as accessibility, can be defined as the potential to spatially interact with opportunities. For decades, transportation planning research and practice have relied on mobility-based estimates and indicators. These have been oriented to reflect realized movement (e.g., kilometers traveled, emissions released, vehicles passing through an intersection) as opposed to potential movement (e.g., the number of opportunities that can be reached). However, in recent years, there have been calls to move from mobility-based methods to access-based ones, but a few barriers remain. One significant issue is methodological--namely, the lack of clarity in the interpretation of conventional accessibility measures' scores. An emerging approach is to link these scores to outcomes, but this thesis proposes a preceding step: clarifying the units of accessibility.

  In this spirit, this thesis is organized into six chapters. First, a review of how accessibility literature largely diverged from the spatial interaction literature is detailed. This lays the foundation for how the addition of proportionality constants that return the units to the measure and balance them to reflect known constraints in the system may be useful. Second, using a synthetic example, the total constrained accessibility measure is introduced. This measure is formally proportional to unconstrained accessibility (i.e., Hansen-type accessibility measure commonly used in literature) but with results consistently expressed in units of opportunities. Next, using the same synthetic example, the single constrained accessibility measure is introduced. This measure considers competition for opportunities and could be understood as related to the popular measures of competitive accessibility. Multimodal extensions as well as the expression of these measure in terms of ‘market potential’--the accessible population at a zone--is also introduced. Third, motivation, data and methods of an empirical example of the parkland area and population in the City of Toronto are detailed. All members of the family of accessibility measures introduced in this work are compared and contrasted in the subsequent chapters using this empirical example. Fourth, a unimodal case considering a binary impedance function and all population as walking is presented for accessible parkland area and accessible population in Toronto. Differences in the unconstrained (i.e., conventional method) and constrained measures are detailed. Fifth, a multimodal case of accessible parkland area and accessible population is detailed, considering walking, cycling, transit, and car modes. Similarly, differences in the unconstrained and constrained measures are detailed. Sixth, this final chapter summarizes key points across all chapters, provides a thorough discussion on how constrained accessibility can be used to communicate for the purpose of equitable transportation policy, and details future lines of potential investigation.
  
# Note that a tab is needed on the line after the `|`. Subsequent paragraphs need to be indented too.
acknowledgements: |
  Oh the places this dissertation has taken me.
  
  First and foremost, I want to express gratitude to my PhD supervisor Antonio Paez, for your unwavering and steady guidance throughout this journey. From co-working coding and writing sessions and weekly check-ins to countless opportunities to present my work at regional and international venues, the freedom and support you've given me to develop ideas, my academic voice and establish my own network over the past four years have made the completition of this dissertation possible. I also want to sincerly thank Moataz Mohamed, my MASc supervisor, who first introduced me to the academic world. Without you, these X words, my X 'academic' words, would not be written. Also, I want to also thank Christopher Higgins, who worked alongside Antonio, Moataz and myself on one of the earliest projects of my PhD journey--work that ultimately laid the foundation for this dissertation.

  I also want to express gratitude for the Mobilizing Justice Partnership, led by Steven Farber and supported academically in its early years by Ignacio Tiznado-Aitken and Matthew Palm--thank you for the guidance and support in framing my academic understanding of transportation equity and justice. I'm also especially grateful for the Partnerships'connection with the University of Toronto and the opportunity to work with others with similar synergies; Joao, Robert, Madeline. Also, for the Partnership's network of transportation-oriented practioners deeply passionate about equity. This network led met to sit alongside the work being done at the City of Toronto facilitated development of the Transportation Equity Policy Framework led by Lorina Hoxina. Serving on the technical advisory committee, I witnessed the dedication of civil servants pushing the boundaries of how equity analysis can be applied in practice. I offered my feedback and advice on methods as appropriate in different forums, and have some lively dicussions with accessibility fanaticis Bryce Sherman, Herman, and Michael. I also ultimately left inspired to apply the methods introduced to a partical example of parkland area they were considering, so impence gratude to them.

  And also to those at McMaster University, I am particularly grateful to Lea Ravensenbergen, whose mentorship opened the door for me to co-supervise my first (and subsequent) undergraduate students. I also want to thank the growing dommunity of transportation reseacher, the Friends of Transportation Fest, Darren Scott and his students Y, S, and Sadia, along with my academic siblings Bruno and Amanada and Javier and Carlos, and the comraderie of Lea and Moataz's students particuarly: Ally, Dastar, X, and Sadia, Ali and Foda. Though I completed the PhD program and this dissertation while living in Toronto, you all made visiting campus worthwhile --rewarding, enriching, and fun. Your perspectives sharpened my thinking and helped me see familiar problems in new ways.

  To my international friends and collavoborate. Namely, to Julio Soria Lara and Alberto, Raul, Manuel, Manu, Amor and other lab mates at TRANSyT at the Universidad de Poltechnic de Madrdid who made my two spring research stays in Madrid insightful, productive, and increadibly unfortable. To the friends, colleagues, and co-conspirators I’ve met at international conferences--NARSC, AAG, TRB, WCTRS, and NECTAR--thank you for the critical conversations and enduring friendships. To the others co-orgainizers of the Transportation Justice group at AAG and the NECTAR Board, for creating spaces where equity and rigor go hand in hand. These international connections made the pursuit of a PhD, so worth it.

  Finally, to my partner in life, and recently my husband, Aidan: thank you for your patience, love, and unwavering belief in me and in humanity’s silly little pursuit of knowledge for a better world. From encouraging me to take the plunge into PhD life, to supporting my obsessive working sessions late into nights before (often self-imposed) deadlines, to standing by me through weeks and months apart while I attended conferences or research stays--your support has been constant. Your belief in me and in the value of continuing this academic journey means more than I can ever fully express.

  This work stands on the shoulders of many. I thank all who came before me--for charting the course I chose and was guided into taking up, allowing me to inch just a bit further.

declarations: |
  I did most of the research.
  
  Also the writing. 
  
  Sometimes I cried.
  
  But mostly I had fun.
quotation: 'SOME QUOTE'
# Author of the quotation
qauthor: 'Some author of a quote'
# Specify the location where the bibliography file is found below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include = FALSE}
# This chunk ensures that the macdown package is
# installed and loaded. This macdown package includes
# the template files for the thesis.
if(!require(remotes))
  install.packages("remotes", repos = "http://cran.rstudio.com")
if(!require(macdown))
  remotes::install_github("paezha/macdown")
library(macdown)
```

<!--
The acknowledgements, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
-->

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Preface - Introduction {.unnumbered}

## This dissertation's evolution

This dissertation has come together non-chronologically. In the summer of 2020, a largely 'virtual' summer for many of us due to the COVID-19 Pandemic declaration and the resulting public health measures, I began working with Dr. Antonio Paez on a reading course as a Masters of Applied Science student under the supervision of Dr. Moataz Mohammed. At that point, I was nearing the end of my Master's program, writing a MASc. thesis aimed at providing guidance on right-sizing passenger transportation sustainability policies based on vehicle technologies well-to-wheel life cycle operating conditions e.g., [@soukhovOccupancyGHGEmissions2022; @soukhovElectricMobilityEmission2022]. Coming from transportation engineering, I had to take on a new research process in engaging with this new research. Antonio, who had recently developed a deep appreciation for open and reproducible science and the R universe, combined with a passion for historical analysis and use of data for advocacy, had worked with students before me to gather information on public schools in Hamilton. A public school in his neighborhood, along with several others across the city, was closing, and he was keen to explore the impacts these closures would have.

It was within this context that I began the reading course. I found myself intuitively working through the available data, building upon my beginner R skills, estimating additional data to supplement the analysis, and crafting research questions that felt sufficiently satisfying; questions that blended policy impacts, spatial analysis, and active transportation implications of a change in school-seats. Spatial accessibility, or the potential to spatially interact with school-seats, as an indicator that could answer these questions. Accessibility, conventionally defined as the sum of opportunities $O_j$ weighted by the travel impedance function $f(c_{ij})$ of some travel cost $c_{ij}$ for a set of zones of origin $i$ $i...I$ and destination $j$ $j...J$ in Equation \ref{eq:access-unconstrained}, takes the following general form:

```{=tex}
\begin{equation}
\label{eq:access-unconstrained}
S_i = \sum_i O_j f(c_{ij}) 
\end{equation} 
```

I wanted as precise a solution possible with an interpretable meaning at the highest spatial resolution. For instance: how had the number of potentially reachable school-seats change for households? Reflecting on this thought process, I placed a high expectation on a place-based accessibility metric. This expectation was motivated by my engineering background that was informing this desire and fixation on units. I knew the number of school-seats in Hamilton for each study year, the number of students, and I knew that, intuitively, the school-seat accessibility for each household should decrease within proximity to schools. However, conventional methods produced results that were difficult to interpret. Moreover, competitive measures, such as those often applied using Floating Catchment Areas (FCA), had their own inflationary issues [i.e., discussed in @paezDemandLevelService2019]. 

This challenge led towards developing a more intuitive approach, one that involved establishing proportional allocation factors. These factors allocate opportunities proportionally based on a zone’s population relative to the region’s total population, as well as the zone's travel impedance to other zones relative to the travel impedance to all zones from that zone. The balancing factors ensure that each parcel is assigned a portion of the total school-seats in the region. This intuitive approached ensured the school-seats from each school is proportionally allocated to each zone based on the zone's relative population and travel impedance _and_ each value summed to the total number of opportunities in the region. This method of competitive accessibility, which preserves its units, was introduced as “spatial availability" in @soukhovIntroducingSpatialAvailability2023 a work co-authored with Chris Higgins, Moataz Mohamed and Antonio Paez.

At the time, we recognized that "spatial availability" is _singly-constrained_ from the perspective of opportunities, akin to the singly-constrained spatial interaction model in @wilson1971. We adopted this conceptualisation, without scrutinizing the Wilson's framework (at the time, despite Chris Higgin's comment "hey, isn't this Wilson's balancing factor?) published "spatial availability", a singly-constrained competitive accessibility measure in _PLOS ONE_ [@soukhovIntroducingSpatialAvailability2023]. An open data paper demonstrating my growing expertise in R and reproducible methods in _Environment and Planning B: Urban Analytics and City Science_ [@soukhovTTS2016RDataSet2023] then followed, followed by a multimodal extension of the spatial availability measure the year after [@soukhovMultimodalSpatialAvailability2024]. Along with years of thinking about the equity and justice implications of transportation systems summarized in the (forthcoming) review of the literature @soukhovSearchingFairnessStandards2025, the first work--that started back in my MASc during that reading course--the article assessing the change in active (and motorized) school-seat accessibility after 10 years of school closures was published in _Networks and Spatial Economics_ [@soukhovTenYearsSchool2025].

However, in the past year, I returned to more deeply inspect the connection with the singly-constrained spatial interaction model and the proportional allocation factor of spatial availability. It turned out: spatial availability can be simplified to be equivalent to the singly-constrained spatial interaction model. Meaning that the most popular competitive accessibility measure--the 2SFCA--can also be seen as proportional to the singly-constrained expression in Wilson (NOTE: in @soukhovIntroducingSpatialAvailability2023, the mathematical equivalence between spatial availability per capita and 2SFCA is demonstrated). And, requiring only information about the total number of opportunities in the region and non-competitive in its definition, a 'total constraint', fitting within Wilson's framework, can also be established. The total constraint ensures the zonal accessibility values are proportional to the total in the region, share the same scaled magnitude as conventional accessibility in Equation \ref{eq:access-unconstrained}, but maintains units of the accessibility concept, i.e. the _number of accessible opportunities_.

Thematically, this thesis begins with this forthcoming (and submitted) work of @soukhovfamilyofaccessibility2025, which details the precedent, logic, and intuition of weaving 'constraints' from the spatial interaction model into measures of potential for spatial interaction: both the singly-constrained version (as previously published) and, for the first time, introducing the total constraint are included.

In summary, this this dissertation is based on the following publications (listed in chronological order of their publication):

- Soukhov, A., Pereira, R. H. M., Higgins, C. H. & Paez, A. (2025). A family of accessibility measures derived from spatial interaction principles. (Forthcoming - submitted to PLOS ONE).
- Soukhov, A., Higgins, C. D., Páez, A., & Mohamed, M. (2025). Ten Years of School Closures and Consolidations in Hamilton, Canada and the Impact on Multimodal Accessibility. Networks and Spatial Economics. https://doi.org/10.1007/s11067-025-09677-z
- Soukhov, A., Aitken, I. T., Palm, M., Farber, S., & Paez, A. (2025). Searching for fairness standards in the transportation literature. (Forthcoming - submitted to *Transportation*). An extended version, as a report, can be found here: https://mobilizingjustice.ca/wp-content/uploads/2024/01/just-transportation-1.pdf 
- Soukhov, A., Tarriño-Ortiz, J., Soria-Lara, J. A., & Páez, A. (2024). Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes. PLOS ONE, 19(2), e0299077. https://doi.org/10.1371/journal.pone.0299077
- Soukhov, A., & Páez, A. (2023). TTS2016R: A data set to study population and employment patterns from the 2016 Transportation Tomorrow Survey in the Greater Golden Horseshoe area, Ontario, Canada. Environment and Planning B: Urban Analytics and City Science, 23998083221146781. https://doi.org/10.1177/23998083221146781
- Soukhov, A., Paez, A., Higgins, C. D., & Mohamed, M. (2023). Introducing spatial availability, a singly-constrained measure of competitive accessibility. PLOS ONE, 1–30. https:// doi.org/10.1371/journal.pone.0278468

Other accessibility research done alongside the completion of this dissertation:
- lea
- lea 2
- school of cities report

## The importance of accessibility; and conceptual issues interpreting 'unconstrained' access

Transportation is a special kind of land-use that compresses space and time, enabling connections, and hence potential interaction, between populations in different zones. But, as posed in @handy2020, "is accessibility an idea whose time has finally come?" This concept has been discussed in the literature for decades as the ulterior goal of transportation systems. Accessibility doesn't express mobility directly but instead captures the potential for meaningful mobility — the reachability of opportunities. Despite the utility of accessibility measures in both literature and practice to identify areas of high and low opportunity access, the interpretability of foundational accessibility measures, such as the Hansen-type accessibility measure [@hansen1959] (as in Equation \ref{eq:access-unconstrained} ), can be challenging [@geurs2004; @millerAccessibilityMeasurementApplication2018; @santanapalacios2022].

By definition, accessibility is the product of the sum of opportunities that can be reached through some travel impedance, meaning each zonal value is sensitive to both inputs in a way that isn't constrained by any known system. For example, consider comparing the accessibility values for a region between two years where the number of opportunities remained constant but the travel behaviour dramatically changed. In this case, both years would have travel impedance functions of different forms. As a result, the zone values would have different units, one in opportunities-weighted-travel-impedance-value-from-form-1 and another in units of opportunities-weighted-travel-impedance-value-from-form-2. These values would exist within different ranges, and while they could be normalized and compared, this process introduces bias. Consider also, when comparing values between regions, the number of opportunities varies with the size of the region (e.g., larger regions generally have more opportunities than smaller ones). For this reason, raw accessibility scores aren't necessarily comparable without some form of normalization, which again introduces bias. As with many spatial issues, these problems are compounded by the Modifiable Areal Unit Problem (MAUP) and Modifiable Temporal Unit Problem (MTUP), where the zonal and temporal units may differ between comparison periods, further obscuring the results.

Aside from temporal and regional comparisons, gleaning interpretation of raw 'opportunities-weighted-travel-impedance-value' scores within the same temporal and spatial region is also challenging. To move closer to an intuitive interpretation, a balancing factor that is sensitive to the known information about the transportation system can be introduced. This would allow the zonal value of 'how many opportunities can be accessed' to be more plainly contemplated - as the units of travel impedance are set aside. In this way, linking accessibility values to something else meaningful (e.g., opportunities or opportunity seekers) become more straightforward and interpretable. This form of intuitive interpretation of the units is a central aspect of this dissertation.

## Aims

The primary contribution of this dissertation is demonstrating how the units of accessibility--the number of opportunities that can potentially be reached--can be preserved within the zonal values of the accessibility measure. This contribution offers a practical way to enhance accessibility as a planning tool, making it more interpretable and potentially more useful for decision-making.

This dissertation offers four key contributions.

First, it defines accessibility as the _potential_ for spatial interaction by explicitly linking it the spatial interaction literature, and providing a review of how the two streams of literature diverged. Next, it explores how spatial interaction modelling literature maintained 'units' in outputs by introducing system and zonal constraints through balancing factors, and shows how accessibility can be reformulated using these same constraints. 

Second, the total constraint is introduced, which is conceptually equivalent in magnitude to the Hansen-type accessibility measure (Equation \ref{eq:access-unconstrained}) but is expressed in units of 'accessible opportunities'. This approach can help ensure that accessibility measures remain consistent and comparable across different areas and time periods, addressing the issue of unit discrepancies that often arise in accessibility analyses.

Third, the single constraint is introduced, which accounts for competition among populations for access to opportunities as well as ensuring the total constraint is met too. The singly-constrained measure can be understood as a competitive measure, and offering an additional layer of constraints (e.g., ensuring opportunities at a destination are proportionally distributed to reachable origins based on relative population and travel impedance, as well as maintaining the total constraint). It can also be seen as a more general version of the Two Step Floating Catchment Area (2SFCA) method [@luo2003]. In this way, singly-constrained accessibility can offer a more accurate understanding of localized competition for opportunities.

Fourth, the multi-modal extension of both total- and singly- constrained accessibility is specified. Specifically, these extensions demonstrate how these measures can be adapted to consider different groups with different travel impedance functions (i.e., be it multiple modes of transport or populations with different travel likelihoods) accessing the same opportunities. This multi-modal extension can be used to interpret the accessibility gap between groups, illustrating how many more opportunities can be reached by one group compared to another.

These are four novel contributions that advance accessibility analysis while increasing its practicality and interpretability. 

## A practical case study: Toronto and publically owned and opertated green space
[**NOT DONE**]

To demonstrate the measures presented in the first two chapters, an empirical example is developed.

The magnitude of opportunities are reachable, and the distribution of this value spatially or among people, is often the preoccupation of those who use accessibility measures. An advantage of the _constrained_ accessibility measures, is in their interpretation. They can be used in place of a static metrics of opportunity per zone or opportunity per capita in a zone metric (e.g., the number of jobs located in a census block, or the number of hospital beds per person in a neighbourhood). Instead, constrained accessibility measures yield a value of the number of _accessible_ opportunities per zone or per capita (in a zone), cleanly folding in the assumptions about how the population may interact with that opportunity (transportation-land-use interaction), but still allowing the value itself to be understood like a level of service metric. 

To demonstrate the practical advantage constrained accessibility's interpretation, a case study reporting the accessibility to green space that is owned and operated by the City of Toronto will be detailed. Green space is important to well being... XYZ. Furthermore, the City is undergoing benchmarking of access to different opportunity types. TRANSPORTATION EQUITY POLICY FRAMEWORK development. From this perspective, greenspace is an important public service to benchmark using accessibility, as its use is related to its ease of reach []. 

Green space is an interesting opportunity, as there are a variety of types with different qualities that may impact how important population-competition considerations are. For instance, smaller neighbourhood parks typically only attract people residing in proximity to these parks, and the furniture may get congested - meaning competition may matter, but only locally. However, data reflection the number of people that visit these parks is more likely to be unavailable. On the other hand, large legacy parks with natural attractions may attract people from further distances, so congestion of the park may be more of a pertinent consideration. 

With the aim of showcasing constrained accessibility's practical interpretation, the calculation and comparision between total constrained and singly constrained accessibility will be outlined in this work. Constrained accessibility for each zonal origin will be calculated using the total constraint assuming one travel impedance function and a multi-modal extension, showing how constrained accessibility, expressed in units of accessible opportunities, can be aggregated across different spatial and temporal scales. This aggregation allows for a clearer communication of the policy implications of transportation and land-use decisions, making accessibility measures more communicable for urban planning and decision-making.

## Overview of methods 
[**NOT DONE**]

### Reproducible statement
Firstly, this entire thesis adopts principles of open and reproducible methods; using R and RStudio workflows, everything from the manuscript, code, and the final manuscript is transparent and freely available for all to replicate [X]. 

CITE JGS -- antonio editorial, comber brunston, ariabs-bel (open data products). 

Link to the repo: X (the repo has a landing page, with reproducibility instructions on this page, and read mes in the data-raw). 

### Accessibility metrics

The dissertation begins with the demonstration of the unconstrained accessibility (Equation \ref{eq:access-unconstrained}), then total constrained accessibility, and singly-constrained accessibility. Framing of the constrained and unconstrained accessibility methods in the spatial interaction literature. 

These are outlined in chapters 1 and 2. and summarised in chapter 3. 

### Data sources and travel time estimates

1) Parks - from City of Toronto portal. 

2) Canadian census for Toronto CMA. Population weighted centroids. Census data at the DA. Neighbourhoods.  

3) Empirical travel flows from the 2022 TTS for leisure trips for different modes (as included in the updated version of the TTS2016R package (text from _"TTS2016R: A data set to study population and employment patterns from the 2016 Transportation Tomorrow Survey in the Greater Golden Horseshoe area, Ontario, Canada"_ is included)

4) Using modified OSM Toronto network and GTFS files + the resulting routed r5r travel times for multiple modes. 
	
## Chapters outline
[**NOT DONE**]
This dissertation is divided into six chapters. The first provides the conceptual and historical framework of this thesis as written in the first half of the manuscript titled _"Family of accessibility measures derived from spatial interaction principles"_ submitted to _PLOS ONE_. This chapter outlines the evolution of the spatial interaction literature, how accessibility literature diverged, and introduces the total and single constraints using numeric examples. 

The second chapter outlines the family of accessibility measures (unconstrained, total constrained, singly constrained, and doubly constrained accessibility) using a simple toy example. It is the second part of the manuscript "Family of accessibility measures derived from spatial interaction principles"_ submitted to _PLOS ONE_. As well as additional writing detailing the context for and the formulation of a multimodal extension of each measure, inspired writing from the published _"Multimodal spatial availability: A singly-constrained measure of accessibility considering multiple modes"_ paper. <!-- rephrase: --> also, inspired by the unimodal focus of the published paper _"Introducing spatial availability, a singly-constrained measure of competitive accessibility"_.

After the first two chapters, all work is original. 

The third chapter details the empirical example of parkland accessibility in Toronto developed in chapters four and five of this dissertation. Namely the origins, the different zonal levels of aggregation, qualities about the destinations, and the routing and trip length assumptions are detailed. The unconstrained and constrained accessibility measures used in the chapters four and five are briefly summarised in context with the empirical example. 

The fourth chapter calculates a unimodal example of parkland accessibility and accessible population (to parkland area) for the city of Toronto, using the unconstrained, total and singly constrained accessibility. Results using different measures are compared, discussed and contextualized.  

The fifth chapter builds on the preceding chapter, and extends the example to consider multiple modes: walking, as well as cycling, transit and car. Multimodal accessibility to parkland (and to population from parkland) is detailed. Results are compared to the unimodal example, and compared, discussed and contextualised. Discussion of competitive gains is inspired by the published paper _"Ten Years of School Closures and Consolidations in Hamilton, Canada and the Impact on Multimodal Accessibility"_ .

The sixth chapter concludes the work. the open and reproducibility of methods learned is discussed (text from _"TTS2016R: A data set to study population and employment patterns from the 2016 Transportation Tomorrow Survey in the Greater Golden Horseshoe area, Ontario, Canada"_ is included), how constrained measures can be interpreted, how this interpretation can aid policy makers, and how policy makers can ultimately use this to plan for inequities (some text pulled from _"Searching for fairness standards in the transportation literature"_). Future research directions (?)


<!--

BEGINNING HERE ADD THE DOCUMENTS THAT CONFORM THE MAIN BODY OF THE THESIS

-->

```{r eval=!knitr::is_latex_output(), child=here::here("01-chap1.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("02-chap2.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("03-chap3.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("04-chap4.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("05-chap5.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("06-conclusion.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("99-references.Rmd")}

```
